Critical assessment of Linear-Response Excitations in Molecular system\\ or \\ On the Localization of the Linear-Response excitations in Open Systems

### 2018-01-08 LRDraft.tex
update

% notation for standard math
\renewcommand{\d}{\partial}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\dd}{{\rm d}}
\renewcommand{\r}{{\bf r}}
\newcommand{\x}{{\bf x}}
\newcommand{\eps}{\epsilon}
\newcommand{\bbomega}{{\bar{\bar\omega}}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\intdef}[3]{\int_{#1}^{#2} \dd {#3}}
\newcommand{\intover}[1]{\int_{-\infty}^{+\infty} \dd {#1}}
\newcommand{\sint}{\mathrlap{\displaystyle\int}
\mathrlap{\textstyle\sum}
\phantom{\mathrlap{\displaystyle
\int}\textstyle\sum}}

@@

Let us extend the the former analysis to self-consistent systems, \emph{i.e.} to the case in which the unperturbed Hamiltonian is itself a function of the density.
In this case the perturbing field $\op\Phi$ has also the effect of adding a further perturbation operator to the perturbing Hamiltonian. In the linear response framework, 
this latter quantity can be expressed as the functional derivatives of $\hnot$ with respect to the density kernel (evaluated at $\dm=\dmnot$) times the value of the
induced density modification, that is:
\be\label{Vptspacedef1}
\op V'(t)= 
\int \dd \r \dd \r' \frac{\delta\hnot}{\delta \rho_0(\r,\r')}
\trace{\dm'(t) \ket{\r} \bra{\r'}} \;.
\ee
The Liouville equation for self-consistent systems can be written by adding the contribution of $\op V'$ to $\op\Phi$ in \eqref{LiouvilleRhoDef2}. 
It reads:
\be\lb{LiouvilleRhoSelfDef1}
\ii \d_t \dm'(t) = \Liouvnot \dm'(t) + \commutator{\op V'(t)}{\dmnot} + \commutator{\op \Phi(t)}{\dmnot} \;.
\ee
We point out that the response density operator obeys a self-consistent equation, since the  induced perturbing field $\op V'$ is itself a function of $\dm'$. 
It is possible to formally recast \eqref{LiouvilleRhoSelfDef1} in the same form of \eqref{LiouvilleRhoDef2} by introducing the \emph{coupling}
superoperator:
\be
\coupl = 
\int \dd \r \dd \r' \commutator{\frac{\delta\hnot}{\delta \rho_0(\r,\r')}}{\dmnot}
\trace{\cdot \ket{\r} \bra{\r'}} \;,
\ee
in this way the first commutator in \eqref{LiouvilleRhoSelfDef1} can be written as the formal action of the coupling superoperator to $\dm'$ and \eqref{LiouvilleRhoSelfDef1}
can be written as:
\be\lb{LiouvilleRhoSelfDef2}
\ii \d_t \dm'(t) = \Liouv \dm'(t) + \commutator{\op \Phi(t)}{\dmnot} \;, \qq
\Liouv = \Liouvnot + \coupl \;.
\ee

@@

\vspace{0.5cm}
\emph{Solution for $\coupl = 0$}
\vspace{0.5cm}

\noindent In this case the $\optr{L}_{q\beta}^{p\alpha}$ matrix has only diagonal non-vanishing elements, given by: 
\be
\trace{\decay{\alpha}{p}\Liouv\excite{q}{\beta}} = H_{0\alpha\beta}\delta_{pq} - \eps_p\delta_{pq}\delta(\alpha-\beta) \;, \qq
\trace{\excite{\alpha}{p}\Liouv\decay{q}{\beta}} = \eps_p\delta_{pq}\delta(\alpha-\beta) - H_{0\beta\alpha}\delta_{pq} \;, 
\ee
where $H_{0\alpha\beta} = \bra{\psi(\alpha)}\hnot\ket{\psi(\beta)}$. The eigenvalue conditions \eqref{LiouvMatrixEigenvectorDef1} imply
that the $Y$ and $\tilde Y$ coefficients introduced in \eqref{XYcoeffDef1} vanish so that both $\tket{\chi^E_p}$ and $\bra{\chi^E_p}$ do not
contribute to the excitation operators \eqref{ExctitationOperatorsDef2}. The eigenvalue operator equation for $\op E$ reduces to:
\be
\sum_q\left(\hnot - \eps_q\right)\ket{\phi^E_q}\bra{\psi_q} = E \sum_q \ket{\phi^E_q}\bra{\psi_q} \;,
\ee
and projecting on $\ket{\psi_q}$ provides the eigenvalue condition for the excitonic state $\ket{\phi^E_p}$:
\be
\left(\hnot - \eps_p\right)\ket{\phi^E_p} = E \ket{\phi^E_p} \;.
\ee
In the same way, by using the eigenvalue condition for $\op{\tilde E}$, we can derive an analogous equation for $\tbra{\phi^E_p}$. To achieve this
result we need the expression of the left action of the (unperturbed) Liouville superoperator defined in \eqref{L0sopDef1}, that can be derived by
observing that:
\be
\trace{\op O' (\Liouvnot \op O)} = \trace{(\op O' \Liouvnot) \op O } =
\trace{\commutator{\op O'}{\op H_0} \op O} \quad \Rightarrow \quad \op O \Liouvnot = 
- \Liouvnot \op O \;.
\ee
So, the eigenvalue operator equation for $\op{\tilde E}$ provides:
\be
\sum_q  \ket{\psi_q} \tbra{\phi^E_q} \left( \hnot - \eps_q \right) =
E \sum_q  \ket{\psi_q} \tbra{\phi^E_q} \;,
\ee
so that:
\be
\tbra{\phi^E_p} \left( \hnot - \eps_p \right) = E \tbra{\phi^E_p} \;.
\ee
So we find that the excitonic states are the left and right eigenvectors of $\hnot$ (that actually coincide since we are considering
hermitian Hamiltonians). The energy eigenvalue is shifted by $\eps_p$, that is the energy of the occupied states involved in the 
excitation.

@@

>>>>

% notation for standard math
\renewcommand{\d}{\partial}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\dd}{{\rm d}}
\renewcommand{\r}{{\bf r}}
\newcommand{\br}{\bar{\bf r}}
\newcommand{\x}{{\bf x}}
\newcommand{\eps}{\epsilon}
\newcommand{\bbomega}{{\bar{\bar\omega}}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\intdef}[3]{\int_{#1}^{#2} \dd {#3}}
\newcommand{\intover}[1]{\int_{-\infty}^{+\infty} \dd {#1}}
\newcommand{\sint}{\mathrlap{\displaystyle\int}
\mathrlap{\textstyle\sum}
\phantom{\mathrlap{\displaystyle
\int}\textstyle\sum}}

@@

Let us extend the the former analysis to self-consistent systems, \emph{i.e.} to the case in which the unperturbed Hamiltonian is itself a function of the density.
In this case the perturbing field $\op\Phi$ has also the effect of adding a further perturbation operator to the perturbing Hamiltonian. In the linear response framework, 
this latter quantity can be expressed as the functional derivatives of $\hnot$ with respect to the density kernel (evaluated at $\dm=\dmnot$) times the value of the
induced density modification, that is:
\be\label{Vptspacedef1}
\op V'(t)= 
\int \dd \r \dd \r' \frac{\delta\hnot}{\delta \rho_0(\r,\r')}
\trace{\dm'(t) \ket{\r} \bra{\r'}} \;.
\ee
The Liouville equation for self-consistent systems can be written by adding the contribution of $\op V'$ to $\op\Phi$ in \eqref{LiouvilleRhoDef2}. 
It reads:
\be\lb{LiouvilleRhoSelfDef1}
\ii \d_t \dm'(t) = \Liouvnot \dm'(t) + \commutator{\op V'(t)}{\dmnot} + \commutator{\op \Phi(t)}{\dmnot} \;.
\ee
We point out that the response density operator obeys a self-consistent equation, since the  induced perturbing field $\op V'$ is itself a function of $\dm'$. 
It is possible to formally recast \eqref{LiouvilleRhoSelfDef1} in the same form of \eqref{LiouvilleRhoDef2} by introducing the \emph{coupling}
superoperator:
\be\lb{couplingSopDef1}
\coupl = 
\int \dd \r \dd \r' \commutator{\frac{\delta\hnot}{\delta \rho_0(\r,\r')}}{\dmnot}
\trace{\cdot \ket{\r} \bra{\r'}} \;,
\ee
in this way the first commutator in \eqref{LiouvilleRhoSelfDef1} can be written as the formal action of the coupling superoperator to $\dm'$ and \eqref{LiouvilleRhoSelfDef1}
can be written as:
\be\lb{LiouvilleRhoSelfDef2}
\ii \d_t \dm'(t) = \Liouv \dm'(t) + \commutator{\op \Phi(t)}{\dmnot} \;, \qq
\Liouv = \Liouvnot + \coupl \;.
\ee

@@

\vspace{0.5cm}
\emph{Equations for $\coupl = 0$}
\vspace{0.5cm}

\noindent In this case the $\optr{L}_{q\beta}^{p\alpha}$ matrix has only diagonal non-vanishing elements, given by: 
\be
\trace{\decay{\alpha}{p}\Liouv\excite{q}{\beta}} = H_{0\alpha\beta}\delta_{pq} - \eps_p\delta_{pq}\delta(\alpha-\beta) \;, \qq
\trace{\excite{\alpha}{p}\Liouv\decay{q}{\beta}} = \eps_p\delta_{pq}\delta(\alpha-\beta) - H_{0\beta\alpha}\delta_{pq} \;, 
\ee
where $H_{0\alpha\beta} = \bra{\psi(\alpha)}\hnot\ket{\psi(\beta)}$. The eigenvalue conditions \eqref{LiouvMatrixEigenvectorDef1} imply
that the $Y$ and $\tilde Y$ coefficients introduced in \eqref{XYcoeffDef1} vanish so that both $\tket{\chi^E_p}$ and $\bra{\chi^E_p}$ do not
contribute to the excitation operators \eqref{ExctitationOperatorsDef2}. The eigenvalue operator equation for $\op E$ reduces to:
\be
\sum_q\left(\hnot - \eps_q\right)\ket{\phi^E_q}\bra{\psi_q} = E \sum_q \ket{\phi^E_q}\bra{\psi_q} \;,
\ee
and projecting on $\ket{\psi_q}$ provides the eigenvalue condition for the excitonic state $\ket{\phi^E_p}$:
\be
\left(\hnot - \eps_p\right)\ket{\phi^E_p} = E \ket{\phi^E_p} \;.
\ee
In the same way, by using the eigenvalue condition for $\op{\tilde E}$, we can derive an analogous equation for $\tbra{\phi^E_p}$. To achieve this
result we need the expression of the left action of the (unperturbed) Liouville superoperator defined in \eqref{L0sopDef1}, that can be derived by
observing that:
\be\lb{leftActionLouv0Def1}
\trace{\op O' (\Liouvnot \op O)} = \trace{(\op O' \Liouvnot) \op O } =
\trace{\commutator{\op O'}{\op H_0} \op O} \quad \Rightarrow \quad \op O \Liouvnot = 
- \Liouvnot \op O \;.
\ee
So, the eigenvalue operator equation for $\op{\tilde E}$ provides:
\be
\sum_q  \ket{\psi_q} \tbra{\phi^E_q} \left( \hnot - \eps_q \right) =
E \sum_q  \ket{\psi_q} \tbra{\phi^E_q} \;,
\ee
so that:
\be
\tbra{\phi^E_p} \left( \hnot - \eps_p \right) = E \tbra{\phi^E_p} \;.
\ee
So we find that the excitonic states are the left and right eigenvectors of $\hnot$ (that actually coincide since we are considering
hermitian Hamiltonians). The energy eigenvalue is shifted by $\eps_p$, that is the energy of the occupied states involved in the 
excitation.

@@

\vspace{0.5cm}
\emph{Generic equations for $\coupl \neq 0$}
\vspace{0.5cm}

To explicitly write the generic eigenvalue equations for the excitonic states it is useful to introduce some further formalism to express the 
self-consistent part of the perturbing Hamiltonian \eqref{Vptspacedef1}. We start by expanding $\op V'$ in the coordinate representation, that is:
\be
\op V'(t)= 
\int \dd \br \dd \br' \dd \r \dd \r' \bra{\br'}\frac{\delta\hnot}{\delta \rho_0(\r,\r')}\ket{\br}
\trace{\dm'(t) \ket{\r} \bra{\r'}} \ketbra{\br}{\br'}\;, \nn
\ee
then we introduce the \emph{coupling kernel}, defined as:
\be\lb{couplingKernelDef1}
U\left[\op O; \op O'\right] \equiv  \int \dd \r \dd \r' \trace{\op O \frac{\delta \hnot }{\delta \rho_0(\r,\r')}
} \trace{\op O' \ket{\r} \bra{\r'}} \;,
\ee
using this quantity we can express $\op V'$ as:
\be\lb{Vptspacedef2}
\op V'(t)= 
\int \dd \br \dd \br' U\left[\ketbra{\br}{\br'};\dm'(t)\right] \ketbra{\br}{\br'}\;,
\ee
and consequently the formal action of the coupling superoperator \eqref{couplingSopDef1} can be expressed as: 
\be\lb{couplingSopDef2}
\coupl = 
\int \dd \br \dd \br' \commutator{\ketbra{\br}{\br'}}{\dmnot} U\left[\ketbra{\br}{\br'};\cdot\right] \;.
\ee
Using this formalism we can define the left action of the coupling superoperator, in analogous way of what done in \eqref{leftActionLouv0Def1}, 
that is:
\be\lb{leftActionCouplingDef1}
\trace{\op O' (\coupl \op O)} = \int \dd\br\dd\br'\trace{\ketbra{\br}{\br'}\commutator{\dmnot}{\op O'}}
U\left[\ketbra{\br}{\br'};\op O\right] \doteq \trace{(\op O' \coupl) \op O } \;,
\ee
and so we have:
\be\lb{leftActionCouplingDef2}
(\op O' \coupl) = \int \dd\br\dd\br'\trace{\ketbra{\br}{\br'}\commutator{\dmnot}{\op O'}}
U\left[\ketbra{\br}{\br'};\op O\right] \;.
\ee
